\documentclass[11pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{titling}
\begin{document}
\author{Ankur Dhoot}
\title{CS 471 HW}
\maketitle

\section*{Q1}

\paragraph*{(a)}
LWSN, ELLT, HAAS, SC, WTHR, BRNG, HEAV (PMU found from HEAV)

\paragraph*{(b)}
LWSN, ELLT, WTHR, HEAV, PMU

\paragraph*{(c)}
LWSN, ELLT, SC, HAAS, WTHR, HEAV, BRNG, REC, PMU

\paragraph*{(d)}
LWSN, ELLT, WTHR, HEAV, PMU

\paragraph*{(e)}
LWSN, ELLT, WTHR, HEAV, HAAS, PMU

\section*{Q2}
\paragraph*{(a)}
It very much depends on the structure of the graph. If solutions are fairly shallow, then BFS will likely do well as it expands all nodes at a given depth before going deeper. On the other hand, if solutions are fairly deep, then DFS will likely be better as it explores as deep as possibly before rewinding and starting a new path. 

\paragraph*{(b)}
In a heuristic search, h(n) must underestimate the cost value to the goal, else it's possible that the algorithm will fail. If we always underestimate the cost, then all nodes which have path cost less than the optimal solution cost will be expanded since these are all possible nodes on the solution path. If we overestimate the cost, we might not expand a node that is actually on the solution path.

\paragraph*{(c)}
Example given on slide 50 of lecture 3. The reason a non-consistent heuristic can lead to some problems with graph heuristic search is that it may not be the case that the first time we expand a node, we have found the optimal f-cost to that node. Since graph search keeps a closed list of visited nodes, we only visit nodes once. Thus, if a node is on the optimal path to the solution, but we expand it prematurely (through a route not on the optimal path), then we won't be able to expand it when we are exploring the optimal path. If a heuristic is consistent, this guarantees us that whenever we expand a node, the optimal path to that node has been found. 

\paragraph*{(d)}
We'll prove that consistency implies admissibility by induction on the optimal path length, m. Let $N_{m}$ denote a node with optimal path length of m. Let $N_{*}$ denote the goal node. Let $h^{*}(N_{m})$ denote the cost of the optimal path for node $N_{m}$. 

Base case : m = 1

In this case, $N_{1}$ is an ancestor of the goal node. By consistency, we have h($N_{1}$) $\leq$ c($N_{1}$, $N_{*}$) + h($N_{*}$) = c($N_{1}$, $N_{*}$) = $h^{*}(N_{1})$ implying that a node with optimal path length of 1 satisfies the admissibility criteria.

Induction Step: Suppose true for all nodes with optimal path length $<$ m. 

Then, by consistency, h($N_{m+1}$) $\leq$ c($N_{m+1}$, $N_{m'}$) + h($N_{m'}$) for all successors $N_{m'}$ of $N_{m+1}$. In particular, since $N_{m+1}$ has optimal path of length m + 1, it has a successor, $N_{m}$, on the optimal path of length m. Applying the induction hypothesis, we get  h($N_{m+1}$) $\leq$ c($N_{m+1}$, $N_{m}$) + h($N_{m}$) $\leq$ c($N_{m+1}$, $N_{m}$) + $h^{*}(N_{m})$ = $h^{*}(N_{m+1})$ 

Thus, consistency implies admissibility. 

\paragraph*{(e)}
Preliminary result:

f-costs are nondecreasing along any path: For any successor n' of n, we have that f(n') = g(n') + h(n') = g(n) + c(n, n') + h(n') $\geq$ g(n) + h(n) by the consistency condition

Whenever a node n is expanded, the optimal path (in terms of f-cost) to that node has been found: Suppose not. That is f(n) $>$ $C^{*}$, the optimal path cost to node n. Then there must be a frontier node n' on the optimal path to n. Since f-costs are non-decreasing along any path and n' lies on the optimal path $\Rightarrow$ f(n') $\leq$ $C^{*}$ $<$ f(n) meaning n' would be expanded first. 

We have proved that whenever a node is expanded, the optimal f-cost to that node has been found. When we expand a goal node, the f-cost is just the true path length cost (since h = 0), meaning we have found the optimal path to a goal node. 

\section*{Q3}


 




\end{document}
